Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-05-21 second-draft/pnas-word-order-si.tex
SI

%\usepackage{lineno}
%\linenumbers

@@

\usepackage{footnote}
\makesavenoteenv{tabular}
\makesavenoteenv{table}

\usepackage{siunitx}

@@

@@

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}. %\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's seventeen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

%\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
%Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
%The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
%Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
%Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
%For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
%For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}

\section{Formalizing Comunicative Efficiency}
\subsection{Relation to Other Work}

Here we discuss how our formalization of communicative efficiency relates to other information-theoretic work on language.

Our formalization of efficiency is identical to the function proposed in \cite{ferreri2003least,ferrericancho2007global,futrell2017memory,peloquin2019interactions} as a general efficiency metric for communicative systems, with syntactic structures taking the role of underlying meanings that the speaker wants to convey.
This function is also equivalent to a deterministic version \cite{strouse2017deterministic} of the Information Bottleneck approach, which has been successfully applied to modeling word meaning across different domains \cite{zaslavsky2018efficient, zaslavsky2019semantic}.
Very similar metrics have been proposed in other computational work formalizing Zipf's ideas~\cite{frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge,piantadosi2012communicative}.
%More generally, these approaches instantiate Shannon's Rate Distortion theory.
Common to all these accounts is the idea of maximizing the \key{amount of information} that linguistic forms provide about meanings, while constraining \key{complexity and diversity} of forms; there are some differences in the formalization of complexity.

All these accounts formalize the amount of information provided as the \key{mutual information} between the forms that a listener receives, and the meanings that the speaker intends to communicate to the speaker.
Most previous studies \cite{ferreri2003least, kemp2012kinship,regier2015word,zaslavsky2018efficient} are concerned with the meanings of individual words; they thus take the forms to be individual words, which refer to individual objects and their properties.
We instead study how words are composed into sentences.
In our work, the underlying meanings are the syntactic structures; the forms are the linearized sentences communicated by the speaker to the listener.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The syntactic structure specifies how the meanings of words are combined; therefore, recovering the syntactic structure is a prerequisite to understanding a sentence correctly.

A few models have used other complexity measures that are not explicitly information-theoretic quantities.
In \cite{regier2007color, xu2014numeral, xu2016historical}, the complexity function is the number of different forms; in \cite{kemp2012kinship} it is the difficulty of defining the concepts that are encoded.
The number of forms is not applicable in our case, as languages will typically have infinitely many sentences.
Notably, the models in \cite{regier2007color, xu2016historical} have since been reformulated successfully in the information bottleneck formalism, bringing them even closer to our formalization.

In addition to these models, which quantify the efficiency of communication systems, there is closely related work formalizing the optimal choice of specific utterances in context, such as models of pragmatic reasoning in reference games \cite{frank2012predicting,goodman2013knowledge, kao2014nonliteral}.
In line with the other models, these assume that rational speakers choose utterances to optimize informativity about the referent object, and trade this off with the cost of the utterance, which is partly chosen to be the surprisal of the utterance \cite{bennett2018extremely}.

\subsection{Possible values of $\lambda$}\label{sec:lambda}

In the efficiency objective
\begin{equation}\label{eq:efficiency}
	R_{\textit{Eff}} := R_{\textit{Pars}} + \lambda R_\textit{Pred}
\end{equation}
the value of $\lambda$ is constrained to be in $[0,1)$.
This means, surprisal must be weighted less strongly than parsability.

Greater values of $\lambda$ mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[L(T); T] \leq H[L(T)].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
\begin{equation}
    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
\end{equation}
and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.
In conclusion, $\lambda$ is constrained to be in $[0,1)$, with values closer to $1$ placing similar weights on both predictability and parsability, whereas values closer to $0$ diminish the role of predictability.

In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.
The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
%We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations, see Table REF.
We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).
%Indeed, the same holds for all of the 38 syntactic relations in UD  (Table~\ref{tab:all-predictions-1}-\ref{tab:all-predictions-2})
%We additionally report results with $\lambda = 0$, i.e., placing all weight on parseability.
That is, at $\lambda$ close to its maximal value, the predictions of optimizing the two scoring functions individually add up to the predictions of efficiency optimization.
Small values of $\lambda$ correspond to the case where predictability plays no role, and only parseability is optimized (Table~\ref{tab:all-predictions-1}).
We also verified the robustness of our results to small variations of $\lambda$, by reporting results with $\lambda=1$ in Table~\ref{table:corr-resu}.

%\section{Relation to the Information Bottleneck}
%
%Our efficiency objective (Equation~\ref{eq:efficiency}), to be maximized for word order grammars $L$, is restated below.
%\begin{equation}
%    \label{eq:efficiency-again}
%    J_T(L) = I[L(T); T] - \lambda H[L(T)].
%\end{equation}
%This equation can be seen as a special case of the objective function proposed in \citet{tishby1999information} for general lossy compression, based on rate-distortion theory \citep{cover2006elements,harremoes2007information}. In that theory, the goal is to derive an encoding $\hat{X}=L(X)$ of some variable $X$ which preserves the information in $X$ relevant to some third variable $Y$, while minimizing irrelevant information. The three variables $Y$,$X$, and $\hat{X}$ are assumed to form a Markov chain $Y \rightarrow X \underrightarrow{L} \hat{X}$. Given those variables, the information bottleneck objective to be maximized is given in Eq.~\ref{eq:ib}.
%\begin{equation}
%    \label{eq:ib}
%    J_{X,Y}(L) = I[\hat{X}; Y] - \lambda I[\hat{X}; X].
%\end{equation}
%In this equation, the joint distribution of $X$ and $Y$ is seen as fixed, and the conditional distribution $L=\hat{X}|X$ is optimized.
%
%In our setting we consider a language $L$ expressing trees $T$ as utterances $U$: $T \underrightarrow{L} U$. To make our formulation more parallel to the information bottleneck, we assume the utterances $U$ are passed through a channel from producer to comprehender, and received as $\hat{U}=U$. We have the Markov chain $T \underrightarrow{L} U \rightarrow \hat{U}$, where we are assuming the conditional channel distribution $\hat{U}|U$ is fixed and the conditional distribution $L=U|T$ is to be optimized. Then our objective is parallel to Eq.~\ref{eq:ib}:
%\begin{equation}
%    \label{eq:efficiency-kinda-ib}
%    J_T(L) = I[U;T] - \lambda I[U;\hat{U}]. % TODO should this be I[\hat{U};T]?
%\end{equation}
%Assuming the channel $U \rightarrow \hat{U}$ is noiseless, we have $U=\hat{U}$, and thus $I[U;\hat{U}]=I[U;U] = H[U]$. Then Eq.~\ref{eq:efficiency-kinda-ib} reduces to a form of Eq.~\ref{eq:efficiency-again}:
%\begin{equation*}
%    J_T(L) = I[U;T] - \lambda H[U].
%\end{equation*}
%
%Our efficiency objective is thus closely related to the information bottleneck; the difference is that we are optimizing the mapping from underlying meanings to representations (strings) assuming that these strings will be transmitted in a fixed noiseless channel.
%
%

@@

In Figure~\ref{fig:pareto-per-lang}, we show the predictability-parseability planes for every one the 51 languages, together with Pareto frontiers estimated from optimized grammars.\footnote{Note that in a few languages, the real grammar is at a position slightly \emph{beyond} the estimated Pareto frontier. Due to use of stochastic gradient descent in optimization, there is some amount of noise in the optimization process.}

We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.

Within each language, we estimate the Pareto curve based on the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
%We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).

Figure (REF) in the main paper shows the average of these per-language plots. \textcolor{red}{unify pareto curve computation in main paper}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
\caption{Predictability and parseability of 51 languages. Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars. All data are $z$-scored.}\label{fig:pareto-per-lang}
\end{figure}

In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).
In Table~\ref{fig:pareto-per-lang-stats}, we provide per-language results for the $t$-tests and binomial tests.

\begin{table}
\centering
\small{
\begin{tabular}{l||ll|lll|lll}
Language & Pred. (t) & Parse. (t) & \multicolumn{3}{|c|}{Pred. (Binomial)} & \multicolumn{3}{|c|}{Parseab. (Binomial)} \\ 
&  $p$ & $p$ &  Est. &CI & $p$ & Est. & CI & $p$  \\ \hline \hline
\input{../results/plane/per-language-statistics.tex}
\end{tabular}
}
\caption{Per-language results in Study 1. For each language, we show the following: (1) $p$-values obtained from a one-sided $t$-test, for the null that the mean predictability/parseability of random grammars is at least as high as that of the real grammar. (2) Results from one-sided binomial tests, for the null that the the real grammar is better than at most $50 \%$ of random grammars. In addition to the $p$-value, we report point estimate and 95\% confidence interval for the fraction of random grammars that have values below real grammars.}\label{fig:pareto-per-lang-stats}
\end{table}

@@

We show predictions for the eight correlations on the level of individual languages in Figure~\ref{fig:per-lang}.

@@

Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
We optimize grammars using the same stochastic gradient descent method as we use for efficiency optimization (Section REF).
Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
Indeed, we found that DLM is predicted by efficiency optimization, as discussed in Section REF.
Results in Table~\ref{tab:all-predictions-1} show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.

@@

&	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline

@@

\begin{table*} % TODO figure out away to make this table more aesthetic
	\begin{center}	
\small{
\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
	Relation & Real & DLM & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
\hline
appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\           
lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\          
expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\         
iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\     
vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\

@@

\subsection{Other Experiments}
In Table~\ref{table:corr-resu} we report the results of the simulations we did before the main experiment reported in the main paper (left and center), together with their preregistrations, besides the results from the main experiment (right).
These experiments all had the same setup described in Section (REF), which was fixed before starting simulations; differences are that (1) the first simulation places fully equal weight on parseability and predictability ($\lambda=1.0$), and (2) the final experiment uses three random seeds per grammar.
Results across all three experiments agree; jointly optimizing grammars for parseability and predictability produces all eight correlations.
\textcolor{red}{TODO also the aggregate posterior}

\begin{table}[hbt!]
& $\lambda=1.0$\footnote{\url{https://aspredicted.org/blind.php?x=bg35x7}. Note that we report the locality results as `DLM' in Table REF.} 
& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=8gp2bt}}
& $\lambda=0.9$\footnote{\url{http://aspredicted.org/blind.php?x=ya4qf8}} \\ \hline \hline 
\caption{\textcolor{red}{probably should also report parseability results as lambda=0 here}Results from previous optimization experiments with $\lambda=1.0$ (left) and $\lambda=0.9$ (center), and the main study (right).}\label{table:corr-resu}

%\footnotetext{}
%  final experiment
%  0.9
%

@@

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
%Our method will be applicable more generally to the task of optimizing grammars for a broad class of objective functions.
We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
Previously, \citet{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).
Empirically, we observe that optimizing differentiable ordering grammars for efficiency leads to convergence towards deterministic behavior, allowing us to extract equivalent deterministic grammars as described in Section~\ref{sec:diff-gramm}.

See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates used in this optimization scheme.

@@

In this section, we describe the details of the neural network architectures.
Choices follow standard practice and machine learning and were fixed before evaluating word order properties, and the efficiency of real languages.

@@

\paragraph{Estimating Predictability}
We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
Given the small size of the corpora, this limit is only attained only for few languages.
In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
The model predicts both the next word and its POS tags in each step.
Using POS tags is intended to prevent overfitting on small corpora.

@@

Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
These parameters are shown in Table~\ref{tab:hyperparameters}.
Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (efficiency, predictability, parseability) on the language. We then used this setting for creating optimized word order grammars.

%\textcolor{red}{TODO random seeds for efficiency}

@@

Further, we report results for grammars optimized for dependency length minimization (DLM, see Section~\ref{sec:DLM}).
Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.

>>>>

@@

@@

\textcolor{red}{cite rijkhof etc}

\textcolor{red}{TODO This idea makes strong predictions, and turns out to overpredict correlations (see SI, Section X).}

@@

We base our formalization on the comprehensive study by Dryer \cite{dryer1992greenbergian}.\footnote{Regarding objections by \cite{dunn2011evolved}, we refer to the follow-ups by \cite{levy2011computational}, and \cite{croft2011greenbergian}.}
Greenberg's original study was based on 30 languages; more recently, Dryer \cite{dryer1992greenbergian} documented the word order correlations based on typological data from 625 languages.
\cite{dryer1992greenbergian} formulated these universals as correlations between the order of objects and verbs and the orders of other syntactic relations.
We test our ordering grammars for these correlations by testing whether the coefficients for these syntactic relations have the same sign as the coefficient of the verb-object relation.
Testing correlations is therefore constrained by the degree to which these relations are annotated in UD.
The verb-object relation corresponds to the  \emph{obj} relation defined by UD.
While most of the other relations also correspond to UD relations, some are not annotated reliably.
We were able formalize eleven out of Dryer's seventeen correlations in UD.
Six of these could not be expressed individually in UD, and were collapsed into three coarse-grained correlations:
First, tense/aspect and negative auxiliaries are together represented by the \emph{aux} relation in UD.
Second, the relation between complementizers and adverbial subordinators with their complement clauses is represented by the \emph{mark} relation.
Third, both the verb-PP relation and the relation between adjectives and their standard of comparison is captured by the \emph{obl} relation.

@@

\textcolor{blue}{Greenberg~\cite{greenberg1963universals} stated a total of 45 universals.
Twenty of these concern the structure of individual words (as opposed to word order, which we focus on here), and many of those have been argued to be explained by the ``dual pressures'' idea \cite{haspelmath2006against}.
The other 25 universals concern word order; Dryer \cite{dryer1992greenbergian} reformulated most of these as correlations with verb-object order, finding 16 correlations to likely be universal.
Excluding universals not annotated in available databases, and merging some that can only be tested together we arrived at the eight correlations shown in Table~\ref{table:corr-dryer}.
%Due to data availability, we restricted to 11 universals and further collapsed to the eight correlations shown in Table~\ref{table:corr-dryer}.
Testing efficiency on the excluded correlations would be feasible with additional annotation of data.
For instance, one of these could be tested by labeling the existing databases for the distinction between manner adverbs from other kinds of adverbs.
For another one, it would be necessary to annotate databases for additional languages that have independent plural markers (e.g., Hawaiian).}

@@

\paragraph{Relative Efficiency of Real Grammars}
In the main paper, we tested whether real grammars are more efficient than the mean of random grammars, using a $t$-test.
We also conducted the analysis using a Binomial test (one-sided), testing whether the real gramar is more efficient than the \emph{median} of random grammars, avoiding any distributional assumption on the random grammars.
As before, we used Hochberg's step-up procedure (Note that the tests for different languages are independent, as different random grammars are evaluated for each language), with $\alpha=0.05$.
In this analysis, real grammars improved in parseability for 80\% of languages, in predictability for 69\% of languages, and in either of both in 92\% of languages ($p<0.05$, with Bonferroni correction).

\paragraph{z-Scored Data (Figure (REF))}
We z-transformed on the level of individual languages, normalizing the mean and SD parseability and predictability of the (1) real grammar, (2) the mean of predictability and parseability of all random grammars, (3) the grammar optimized for efficiency (at $\lambda =0.9$, see Section~\ref{sec:lambda}), (4) grammar optimized for parseability only, and (5) grammar optimized for predictability only.

Within each language, we estimate the Pareto curve by connecting the points obtained by optimizing for (1) efficiency (at $\lambda = 0.9$, see Section~\ref{sec:lambda}), (2) parseability, and (3) predictability.
We then estimated the overall average Pareto curve, as shown in Figure (REF), by averaging the z-transformed points (1), (2), (3).

\paragraph{Per-Language Results}
Figure~\ref{fig:pareto-per-lang}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{../results/plane/pareto-plane-perLanguage.pdf}
\caption{per language results. Green: random baselines, Red: real grammar, blue: approximate Pareto frontier, computed from the optimized grammars.\textcolor{red}{would be good to have same colors as in main paper} All data are $z$-transformed.}\label{fig:pareto-per-lang}
\end{figure}

\textcolor{red}{also provide p value, CI for how many baselines are better in either of the two dimensions}

@@

We show results on the level of individual languages in Figure~\ref{fig:per-lang}.

@@

Further, we report results for grammars optimized for dependency length minimization (DLM).
Following \cite{futrell2015largescale}, we formalize the dependency length of a sentence as the sum of the distances of all pairs of syntactically related words, where distance is measured by the number of intervening words.
Prior work has suggested principles akin to DLM as approximating efficiency optimization of grammars~\cite{hawkins1994performance,futrell2017memory, futrell2017generalizing}.
Indeed, we found that DLM is predicted by efficiency optimization, as reported in Discussion (Main Paper).
In line with these ideas, we show that optimizing for DLM makes predictions similar to efficiency optimization, itself a novel result.

@@

&	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline

@@

\begin{table*} % TODO figure out away to make this table more aesthetic
	\begin{center}	
\small{
\begin{tabular}{|l|l|l|ll|l|l|llllllllllllllllllllllllllll}
	\hline
	Relation & Real & DepL & Pred & Pars & Efficiency & Expected Prevalence  \\ \hline
\hline
appos  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_appos.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_appos.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_appos.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_appos.pdf}   &                                               Unknown \\           
lifted\_cc  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_lifted_cc.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_lifted_cc.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_lifted_cc.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_lifted_cc.pdf}   &                  Unknown \\          
expl  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_expl.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_expl.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_expl.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_expl.pdf}   &                                                        Unknown \\         
iobj  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_iobj.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_iobj.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_iobj.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_iobj.pdf}   &                                                        Unknown \\     
vocative  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Real_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_DependencyLength_vocative.pdf}   &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Predictability_vocative.pdf}  &   \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Parseability_vocative.pdf}  &  \includegraphics[width=0.06\textwidth]{../results/correlations/figures/posteriors/posterior_perRelation_Efficiency_vocative.pdf}   &                            Unknown \\

@@

\section{Possible values of $\lambda$}\label{sec:lambda}

In the efficiency objective~(\ref{eq:efficiency-again}), the value of $\lambda$ is constrained to be in $[0,1)$.
This means, surprisal must be weighted less strongly than parsability.

Greater values of $\lambda$ mathematically result in degenerate solutions.
To show this, note that the following inequality always holds:
\begin{equation}
I[L(T); T] \leq H[L(T)].
\end{equation}
Therefore, if $\lambda \geq 1$, the efficiency objective satisfies 
\begin{equation}
    J_T(L) = I[L(T); T] - \lambda H[L(T)] \leq 0.
\end{equation}
and it takes the maximally possible value zero if $L(T)$ maps all trees to a single utterance, in which case both $I[L(T); T]$ and $H[L(T)]$ are zero.
This is a degenerate language with only a single utterance, which is simultaneously used to convey all meanings.
While the design of our word order grammars precludes a collapse of all trees to a single utterance, this shows that an objective with $\lambda \geq 1$ cannot be a generally applicable objective for language efficiency.

In our experiments, we chose $\lambda = 0.9$ as a mathematically valid value that puts similar weight on both predictability and parsability.

The computational cost of grammar optimization precluded repeating the experiment for many values of $\lambda$.
We previously carried out language optimization for $\lambda = 1$, and obtained essentially identical predictions for word order correlations.
We note that each of the eight correlations is predicted by at least parseability or predictability, without any contradictory predictions (Table~\ref{tab:all-predictions-1}).

@@

In this section, we describe the method we employ for creating grammars that are optimized for efficiency.
We carry out optimization in an extended space of grammars that interpolates continuously between different grammars (Section~\ref{sec:diff-gramm}).
More specifically, we include probabilistic relaxations of grammars, which describe probability distributions over different ways of ordering a syntactic structure into a sentence.
This makes efficiency a \emph{differentiable} function of the grammar parameters, and enables efficient optimization with stochastic gradient descent, as we describe in Section~\ref{sec:optim-eff}.

This method addresses a major challenge noted in previous work optimizing grammars, namely that the predictability (and parseability) of an individual sentence depends on the entire distribution of the language.
Previously, \citep{gildea2015human} optimized grammars for dependency length and trigram surprisal using a simple hill-climbing method on the grammar parameters, which required reestimating the trigram surprisal model in every iteration.
Such a method would be computationally prohibitive for efficiency optimization, as it would require reestimating the neural network models after every change to the grammar, which would amount to reestimating them hundreds or thousands of times per grammar.
Our method, by allowing for the use of stochastic gradient descent, addresses this challenge, as we describe in Section~\ref{sec:optim-eff}.

@@

At the beginning of the optimization procedure, we initialize all values $a_\tau := 0.5$, $b_\tau := 0$ (except for the \emph{obj} dependency, for which we fix $a_\tau$ to 0 or 1, see Section~\ref{sec:neural-architectures}).

See Section~\ref{sec:neural-architectures} for the stopping criterion and learning rates.

@@

In this section, we describe the details of the neural network architectures.
All choices described in this section were fixed before evaluating word order properties and the efficiency of real language.

@@

\paragraph{Estimating Predictability}
We choose a standard LSTM \citep{hochreiter1997long} language model with vocabulary size restricted to the most frequent 50,000 words in the treebanks for a given language.
Given the small size of the corpora, this limit is only attained only for few languages.
In each time step, the input is a concatenation of embeddings for the word, for language-specific POS tags, and for universal POS tags.
The model predicts both the next word and its POS tags in each step.
Using POS tags is intended to prevent overfitting on small corpora.
These decisions were made before evaluating word order properties.

@@

Neural network models have hyperparameters such as the number of hidden units, and the learning rate. 
For predictability and parseability optimization, we first selected hyperparameters on the respective objectives for selected languages on the provided development partitions.
Then, for each language and each objective function, we created eight random combinations of these selected hyperparameter values, and selected the setting that yielded the best value of the respective objective function (predictability or parseability) on the language. We then used this setting for creating optimized word order grammars.

\textcolor{red}{TODO random seeds for efficiency}

@@
